Untitled

### 2017-02-02 Manuscript/Manuscript using actual data.Rmd

ggplotly(
  ggplot(subset(cod.df.males, era == "1969-1983"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = COD), 
              color = "white") +
    geom_segment(data = bw_dat.males,
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.males,
                 aes(x = first.gap, xend= first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

@@

ggplotly(
  ggplot(subset(cod.df.males, era == "1983-1993"), aes(y = state.gdiff.num.2, x = new.start.2)) +
    geom_rect(aes(xmin = new.start.2, ymin = state.gdiff.num.2 - 0.45, 
                  ymax = state.gdiff.num.2 + 0.45, 
                  xmax = new.finish.2, fill = COD), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.2 - 0.5, yend = state.gdiff.num.2 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks.2, labels = y.labels.2)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))
```

@@

ggplotly(
  ggplot(subset(cod.df.males, era == "1993-2013"), aes(y = state.gdiff.num.3, x = new.start.3)) +
    geom_rect(aes(xmin = new.start.3, ymin = state.gdiff.num.3 - 0.45, 
                  ymax = state.gdiff.num.3 + 0.45, 
                  xmax = new.finish.3, fill = COD), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = fourth.gap, xend= fourth.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.3 - 0.5, yend = state.gdiff.num.3 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks.3, labels = y.labels.3)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1993-2013"), yaxis = list(title = NA, autorange = "reversed"))
```

@@

>>>>

males.era1 <- ggplotly(
  ggplot(subset(cod.df.males, era == "1969-1983"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = COD), 
              color = "white") +
    geom_segment(data = bw_dat.males,
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.males,
                 aes(x = first.gap, xend= first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

males.era1
```

@@

males.era2 <- ggplotly(
  ggplot(subset(cod.df.males, era == "1983-1993"), aes(y = state.gdiff.num.1, x = new.start.2)) +
    geom_rect(aes(xmin = new.start.2, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.2, fill = COD), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

males.era2 
```

@@

males.era3 <- ggplotly(
  ggplot(subset(cod.df.males, era == "1993-2013"), aes(y = state.gdiff.num.1, x = new.start.3)) +
    geom_rect(aes(xmin = new.start.3, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.3, fill = COD), 
              color = "white") +
    geom_segment(data = bw_dat.males, 
                 aes(x = fourth.gap, xend= fourth.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.males, 
                 aes(x = third.gap, xend= third.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>% 
  layout(xaxis = list(title = "Life expectancy gap (years): 1993-2013"), yaxis = list(title = NA, autorange = "reversed"))

males.era3
```

@@

```{r, fig.align="center", fig.height=5, fig.width=9, echo=F, eval = F}
subplot(males.era1, males.era2, males.era3, shareY = T)
```

**Figure 1a: Decomposition of the change in the female black-white life expectancy gap by cause of death between 1969 and 1983 by US state**

cod.df.females <- merge(subset(cod.df, sex == "Female"), bw_dat.females, by = c("state"))
cod.df.females <- cod.df.females %>% mutate(new.start.1 = -start + first.gap, new.finish.1 = -finish + first.gap,
                                        new.start.2 = -start + second.gap, new.finish.2 = -finish + second.gap,
                                        new.start.3 = -start + third.gap, new.finish.3 = -finish + third.gap)

y.labels <- levels(factor(cod.df.females$state.gdiff.order.1))
y.num.breaks <- length(levels(factor(cod.df.females$state.gdiff.order.1)))

females.era1 <- ggplotly(
  ggplot(subset(cod.df.females, era == "1969-1983"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = COD), 
              color = "white") +
    geom_segment(data = bw_dat.females,
                 aes(x = second.gap, xend= second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) +
    geom_segment(data = bw_dat.females,
                 aes(x = first.gap, xend= first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) +
    theme_minimal() + scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

females.era1
```

- The gap in life expectancy narrowed for all states shown, except for Arizona and small increase in DC.

- CVD was the biggest contributor to the gap narrowing, but all other causes and communicable diseases led to narrowing in many states. Injuries and non-communicable disease also led to narrowing, but the contribution was small. Cancer prevented the gap from narrowing further in some states.
